---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5)
```

This is the report to compare the comparisons between the species datasets.

```{r bind, include = FALSE}
# base_path <- "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>%
  mutate(smoothing_type = gsub("Smoothing", "", smoothing_type))

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")) #, "LSTM"))
data$Species <- factor(data$Species, levels = c(unique(data$Species)))

# save this
#print(head(data %>% dplyr::filter(Metric == "F1")))
fwrite(data, file.path(base_path, "Output", "collated_performance.csv"))
```

## Part One: Base Performance
Looking at base performance.
```{r}
means <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1", smoothing_type == "No") %>%
  group_by(smoothing_type, Species) %>%
  summarise(Score = mean(Score))

print("best performance")
max(means$Score)
print("worst performance")
min(means$Score)

```

## Part Two: Overall Post-processing effect
Firstly I'll look at how much improvement there was overall
```{r}
df <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1")
# and also do it as relative from the control
setDT(df)

# Get NoSmoothing scores by Species
baseline <- df[
  smoothing_type == "No",
  .(Species, rep_id, baseline = Score)
]
df <- merge(
  df,
  baseline,
  by = c("Species", "rep_id"),
  all.x = TRUE
)
df[, relative_change := Score - baseline]
df_sum <- df %>%
  filter(smoothing_type != "No") %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    mean_rc = mean(relative_change, na.rm = TRUE),
    se_rc   = sd(relative_change, na.rm = TRUE) / sqrt(n())
  )

```

```{r compare, fig.width = 8, fig.height = 8}
# its hard to make any one single judgement, so I can make a plot for now

# Plot so that control is more obvious
ggplot(df, aes(x = smoothing_type, y = Score, fill = smoothing_type)) +
  geom_boxplot() +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Activity", y = "Score", fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species)

df <- df %>% dplyr::filter(!smoothing_type == "No")
ggplot(df, aes(x = smoothing_type, y = relative_change, fill = smoothing_type)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Smoothing Type", y = "Relative change to F1 Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species, ncol = 5) +
  theme(legend.position = "none")

```

## Part Four: Within species
Within species
```{r}
# per species
species <- unique(df$Species)
all_stats <- data.frame()
for (x in species){
  dat <- df %>% dplyr::filter(Species == x, 
                              !smoothing_type == "No")
  mod <- lm(
    relative_change ~ 0 + smoothing_type + (1 | rep_id),
    data = dat
  )

  stats <- as.data.frame(summary(mod)$coefficients)
  stats$Species <- x
  stats$sig <- ifelse(stats$`Pr(>|t|)` < 0.05, TRUE, FALSE)
  stats$smoothing_type <- gsub("smoothing_type", "", rownames(stats))
  
  all_stats <- rbind(all_stats, stats)
}

fwrite(all_stats, file.path(base_path, "Output", "WithinSpeciesStats.csv"))

sig_stats <- all_stats %>% dplyr::filter(sig == TRUE)
sig_stats
```

## Part Five: Finding covariates to explain variation
Across species
```{r}
# all species
simple_model <- lmer(relative_change ~ 0 + smoothing_type + (1|Species/rep_id), data = df)
summary(simple_model)
```

Step 3 is to correlate this with the sequential stats I was able to draw out in the sequence report on the training data. Why are some datasets positively affected while others aren't?

```{r extracting stat data}
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv"))

full_df <- merge(df, stat_file, by = "Species")
```

```{r some stats}
# scale everything
full_df <- full_df %>%
  mutate(across(
    c(mean_transitions, total_transitions, Num_Activities),
    ~ scale(.x, center = TRUE, scale = TRUE)[,1]
  ))

full_model <- lmer(
  relative_change ~ 0 +
    smoothing_type *
      (mean_transitions +
       multi_behaviour +
       Num_Activities) +
    (1 + smoothing_type | Species) + 
    (1 | Species:rep_id),
  data = full_df,
  REML = FALSE
)

model_null <- lmer(
  relative_change ~ 0 + smoothing_type +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)

anova(model_null, full_model)


summary(full_model)

library(MuMIn)
r.squaredGLMM(full_model)


```
