---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/oaw001/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 6)
```

This is the report to compare the comparisons between the species datasets.

Step 1 is to pull together all the data.
```{r bind, include = FALSE}
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>% filter(Activity == "Macro-Average")

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("NoSmoothing", "ModeSmoothing", "DurationSmoothing", "ConfusionSmoothing", "TransitionSmoothing", "HMMSmoothing", "BayesianSmoothing", "KalmanSmoothing", "LSTMSmoothing"))

data$Species <- factor(data$Species, levels = c(unique(data$Species)))

```
```{r}
# display this
print(head(data %>% filter(Metric == "F1")))
```

Step 2 is to compare these results. 
```{r compare}
# its hard to make any one single judgement, so I can make a plot for now

# Plot so that control is more obvious
df <- data %>% filter(Activity == "Macro-Average", Metric == "F1")
ggplot(df, aes(x = Species, y = Score, fill = smoothing_type)) +
  geom_point(
    data = df[df$smoothing_type != "NoSmoothing", ],
    aes(x = Species, y = Score, fill = smoothing_type),
    shape = 21,
    colour = "white",
    size = 3
  ) +
  geom_point(
    data = df[df$smoothing_type == "NoSmoothing", ],
    aes(x = Species, y = Score),
    shape = 8,
    colour = "black",
    size = 3
  ) +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Activity", y = "Score", fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# and also do it as relative from the control
setDT(df)

# Get NoSmoothing scores by Species
baseline <- df[smoothing_type == "NoSmoothing", .(Species, baseline = Score)]
df <- merge(df, baseline, by = "Species", all.x = TRUE)
# Calculate relative change from NoSmoothing
df[, relative_change := (Score - baseline) / baseline]
# now plot that
ggplot(df, aes(x = Species, y = relative_change, fill = smoothing_type)) +
  geom_point(
    data = df[df$smoothing_type != "NoSmoothing", ],
    aes(x = Species, y = relative_change, fill = smoothing_type),
    shape = 21,
    colour = "white",
    size = 3
  ) +
  geom_point(
    data = df[df$smoothing_type == "NoSmoothing", ],
    aes(x = Species, y = relative_change),
    shape = 8,
    colour = "black",
    size = 3
  ) +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Activity", y = "Relative change to F1 Score", fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Step 3 is to correlate this with the sequential stats I was able to draw out in the sequence report on the training data.

```{r extracting stat data}
stat_files <- list.files(file.path(base_path, "Output"), pattern = "Sequence_stats.csv", full.names = TRUE, recursive = TRUE)

stat_data <- lapply(stat_files, function(x){
  data <- fread(x)
  return(data)
})

stat_data <- bind_rows(stat_data)
head(stat_data)
```

Of these stats, the ones I think will be most predictive are Prop_Transitions (which represents how many of the training sequences contained transitions) and Transition_Rate (which is how quickly the behaviours turnover within those sequences). Therefore:

```{r some stats}
# add the 2 information sources together
full_df <- merge(df, stat_data, by = "Species")

lm <- lm(relative_change ~ smoothing_type + Mean_Seq_Length + Transition_Rate + Prop_Transitions, data = full_df)
summary(lm)


```
