---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5)
```

This is the report to compare the comparisons between the species datasets.

```{r bind, include = FALSE}
# base_path <- "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>%
  mutate(smoothing_type = gsub("Smoothing", "", smoothing_type))

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")) #, "LSTM"))
data$Species <- factor(data$Species, levels = c(unique(data$Species)))

# save this
#print(head(data %>% dplyr::filter(Metric == "F1")))
fwrite(data, file.path(base_path, "Output", "collated_performance.csv"))
```

## Part One: Base Performance
Looking at base performance.
```{r}
means <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1", smoothing_type == "No") %>%
  group_by(smoothing_type, Species) %>%
  summarise(Score = mean(Score))

print("best performance")
max(means$Score)
print("worst performance")
min(means$Score)

```

## Part Two: Overall Post-processing effect
Firstly I'll look at how much improvement there was overall
```{r}
df <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1")
# and also do it as relative from the control
setDT(df)

# Get NoSmoothing scores by Species
df[, rep_id := seq_len(.N), by = .(Species, smoothing_type)]
baseline <- df[
  smoothing_type == "No",
  .(Species, rep_id, baseline = Score)
]
df <- merge(
  df,
  baseline,
  by = c("Species", "rep_id"),
  all.x = TRUE
)
df[, relative_change := Score - baseline]
df_sum <- df %>%
  filter(smoothing_type != "No") %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    mean_rc = mean(relative_change, na.rm = TRUE),
    se_rc   = sd(relative_change, na.rm = TRUE) / sqrt(n())
  )

print("number of datasets that were positively influenced by at least one method")
df_sum %>%
  group_by(Species) %>%
  summarise(any_positive = any(mean_rc > 0, na.rm = TRUE)) %>%
  summarise(
    n_species_positive = sum(any_positive),
    n_species_total = n()
  )
print("number of methods that had positively influence in at least one dataset")
df_sum %>%
  group_by(smoothing_type) %>%
  summarise(any_positive = any(mean_rc > 0, na.rm = TRUE)) %>%
  summarise(
    n_smoothing_types_positive = sum(any_positive),
    n_smoothing_types_total = n()
  )

```

```{r compare, fig.width = 8, fig.height = 8}
# its hard to make any one single judgement, so I can make a plot for now

# Plot so that control is more obvious
ggplot(df, aes(x = smoothing_type, y = Score, fill = smoothing_type)) +
  geom_boxplot() +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Activity", y = "Score", fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species)

df <- df %>% dplyr::filter(!smoothing_type == "No")
ggplot(df, aes(x = smoothing_type, y = relative_change, fill = smoothing_type)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Smoothing Type", y = "Relative change to F1 Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species, ncol = 5) +
  theme(legend.position = "none")

```

## Part Three: Per behavioural breakdown
Per behaviour
```{r fig.width = 20, fig.height = 15}
# reading in the behaviour duration stats
durationfiles <- list.files(file.path(base_path, "Output"), pattern = "Duration_stats.csv", full.names = TRUE, recursive = TRUE)
durationdata <- lapply(durationfiles, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  data$Activity <- as.character(data$Activity)
  return(data)
})
durationdata <- bind_rows(durationdata)

plot_data <- data %>%
  filter(Activity != "Macro-Average",
         Metric == "F1")

# merge together
plot_data <- merge(plot_data, durationdata, by = c("Species", "Activity"))

# define replicate id
plot_data[, rep_id := seq_len(.N), by = .(Species, smoothing_type, Activity)]

# extract baseline (No smoothing)
baseline <- plot_data[
  smoothing_type == "No",
  .(Species, Activity, rep_id, baseline = Score)
]

# merge baseline
plot_data <- merge(
  plot_data,
  baseline,
  by = c("Species", "Activity", "rep_id"),
  all.x = TRUE
)

# compute relative change
plot_data[, relative_change := Score - baseline]

ggplot(
  plot_data %>% filter(smoothing_type != "No"),
  aes(x = as.factor(Activity), y = relative_change, fill = smoothing_type)
) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(
    x = "Activity",
    y = "Relative change vs no smoothing",
    fill = "Smoothing Type"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Species, scales = "free")
# 
# mod <- lmer(relative_change ~ mean_c * smoothing_type + (1|rep_id), data = plot_data)
# summary(mod)
# 
# ggplot(plot_data, aes(x = mean_c, y = relative_change)) + geom_point() +
#     facet_wrap(~Species, ncol = 5) +
#   theme(legend.position = "none")

```

## Part Four: Within species
Within species
```{r}
# per species
species <- unique(df$Species)
all_stats <- data.frame()
for (x in species){
  dat <- df %>% dplyr::filter(Species == x, 
                              !smoothing_type == "No")
  mod <- lm(
    relative_change ~ 0 + smoothing_type + (1 | rep_id),
    data = dat
  )

  stats <- as.data.frame(summary(mod)$coefficients)
  stats$Species <- x
  stats$sig <- ifelse(stats$`Pr(>|t|)` < 0.05, TRUE, FALSE)
  stats$smoothing_type <- gsub("smoothing_type", "", rownames(stats))
  
  all_stats <- rbind(all_stats, stats)
}

fwrite(all_stats, file.path(base_path, "Output", "WithinSpeciesStats.csv"))

sig_stats <- all_stats %>% dplyr::filter(sig == TRUE)
sig_stats
```

## Part Five: Finding covariates to explain variation
Across species
```{r}
# all species
simple_model <- lmer(relative_change ~ 0 + smoothing_type + (1|Species/rep_id), data = df)
summary(simple_model)
```

Step 3 is to correlate this with the sequential stats I was able to draw out in the sequence report on the training data. Why are some datasets positively affected while others aren't?

```{r extracting stat data}
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv"))

full_df <- merge(df, stat_file, by = "Species")
```

```{r some stats}

full_model <- lmer(
  relative_change ~ 0 + smoothing_type * transition_rate * multi_behaviour * Num_Activities * total_transitions +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)
#summary(full_model)

simple_model <- lmer(
  relative_change ~ 0 + smoothing_type * transition_rate +
    smoothing_type * multi_behaviour +
    smoothing_type * total_transitions +
    smoothing_type * Num_Activities +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)
#summary(simple_model)

very_simple_model <- lmer(
  relative_change ~ 0 + smoothing_type + transition_rate + multi_behaviour + total_transitions + Num_Activities +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)
# summary(very_simple_model)

anova(full_model, simple_model, very_simple_model)

sig_terms <- summary(simple_model)$coefficients %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(`Pr(>|t|)` < 0.05)
sig_terms

```
