---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5)
```

This is the report to compare the comparisons between the species datasets.

```{r bind, include = FALSE}
# Firstly, loading in the individual files and creating the collated performance file

# base_path <- "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>%
  mutate(smoothing_type = gsub("Smoothing", "", smoothing_type))

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")) #, "LSTM"))
data$Species <- factor(data$Species, levels = c(unique(data$Species)))

# save this
#print(head(data %>% dplyr::filter(Metric == "F1")))
fwrite(data, file.path(base_path, "Output", "collated_performance.csv"))
```

## Base Performance
Looking at base performance.
```{r}
means <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1", smoothing_type == "No") %>%
  group_by(smoothing_type, Species) %>%
  summarise(Score = mean(Score))

print("best performance")
max(means$Score)
print("worst performance")
min(means$Score)
print("mean performance")
mean(means$Score)

```

## Overall effect
Firstly I'll look at how much improvement there was overall (across all datasets)
```{r Stats on the overall effect}
# have added in all of Dave's stats corrections. Thank you Dave!!!! Explanatory comments mostly his below.

data <- fread(file.path(base_path, "Output", "collated_performance.csv")) # performance
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv")) # stats about the training data
full_df <- merge(data, stat_file, by = "Species")

df <- full_df %>% filter(Activity == "Macro-Average", Metric == "F1")
baseline <- df %>%
  filter(smoothing_type == "No") %>%
  select(Species, rep_id, baseline = Score)
df <- df %>%
  left_join(baseline, by = c("Species", "rep_id")) %>%
  mutate(relative_change = Score - baseline,
         sp_rep = str_c(Species, "_", rep_id), #*** Reusing factor levels in random effects can cause issues, so I create unique factor levels...
         smoothing_type = factor(smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")), #*** do this in the code flow...
         across(where(is.character), as.factor)) #*** I'd make everything that is still character into a factor, too

# remove no
df1 <- filter(df, smoothing_type != "No") %>%  #*** New df without "No"
  mutate(smoothing_type = fct_drop(smoothing_type)) # Drop unused factor levels

m1 <- glmmTMB(relative_change ~ smoothing_type + (1|Species/sp_rep), data = df1)
summary(m1) #*** `Mode` is now the base case, and all other estimates are adjustments relative to that. To get adjustments relative to zero, just remove the intercept.
m1.2 <- update(m1, ~. -1) #*** the `-1` removes the intercept
summary(m1.2)

```
And a plot that shows the main effects averaged over the random effects.
```{r fig.width = 8, fig.height = 4}
DHARMa::simulateResiduals(m1.2, plot = TRUE) #*** The diagnostics are horrible, at least for Normality...I don't know how we could fix those, though...given the data we have...maybe we just rely on the robustness of the method

#*** If you wanted an overall plot that illustrates the results for the fixed effects averaged over the random effects (mirroring your summary table)
library(marginaleffects)
group_preds <- predictions(m1.2,
                           re.form = NA) %>%   # Includes fixed + random effects
  as_tibble()

overall_plot <- ggplot(group_preds,
       aes(x = smoothing_type,
           y = estimate,
           colour = smoothing_type)) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = .05) +
  geom_point(size = 5) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             colour = "red") +
  labs(x = "Smoothing Type",
       y = "Mean change in F1 (relative to No smoothing)") +
  my_theme() +
  scale_colour_manual(values = my_colours) +
  theme(legend.position = "none")

overall_plot
ggsave(file.path(base_path, "Output", "Overall_effects.png"), overall_plot,
       width = 8, height = 4, units = "in")
```

And looking at the per-species breakdown?

```{r plots on the overall effect, fig.width = 8, fig.height = 8}
# Summarise mean and 95% CI
summary_df <- df1 %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    n = n(),
    mean_rel = mean(relative_change, na.rm = TRUE),
    se = sd(relative_change, na.rm = TRUE) / sqrt(n),
    ci = 1.96 * se, # gives us the confidence intervals (i think)
    .groups = "drop"
  )
# this is 0 every time, so replace with a flat line for visually cleaner
species_plot <- ggplot(summary_df %>% dplyr::filter(!smoothing_type == "No"),
       aes(x = smoothing_type,
           y = mean_rel,
           colour = smoothing_type)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_rel - ci,
                    ymax = mean_rel + ci),
                width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_colour_manual(values = my_colours) +
  labs(x = "Smoothing Type",
       y = "Mean change in F1 (relative to No smoothing)",
       fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  facet_wrap(~Species, ncol = 5)

species_plot
ggsave(file.path(base_path, "Output", "Species_effects.png"), species_plot,
       width = 8, height = 8, units = "in")
```

And expressing these as % performance changes instead...
```{r}
percent_changes <- df %>%
  mutate(percent_change = (relative_change * 100) / baseline) %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    n = n(),
    mean_percent_change = mean(percent_change, na.rm = TRUE),
    se = sd(percent_change, na.rm = TRUE) / sqrt(n),
    ci = 1.96 * se, # gives us the confidence intervals (i think)
    .groups = "drop"
  )

species_plot2 <- ggplot(percent_changes %>% dplyr::filter(smoothing_type != "No"),
       aes(x = smoothing_type,
           y = mean_percent_change,
           colour = smoothing_type)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_percent_change - ci,
                    ymax = mean_percent_change + ci),
                width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_colour_manual(values = my_colours) +
  labs(x = "Smoothing Type",
       y = "Percentage change to the F1 Score",
       fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  facet_wrap(~Species, ncol = 5)

species_plot2
ggsave(file.path(base_path, "Output", "Species_effects_percent.png"), species_plot2,
       width = 8, height = 8, units = "in")


max(percent_changes$mean_percent_change)
min(percent_changes$mean_percent_change)
percent_changes %>% dplyr::filter(smoothing_type == "Bayesian") %>% summarise(mean(mean_percent_change))


percent_changes1 <- df %>%
  mutate(percent_change = (relative_change * 100) / baseline) %>%
  dplyr::filter(smoothing_type != "No") %>%  
  mutate(smoothing_type = fct_drop(smoothing_type))

m3 <- glmmTMB(percent_change ~ smoothing_type + (1|Species/sp_rep), data = percent_changes1)
summary(m3)
m3.2 <- update(m3, ~. -1) #*** the `-1` removes the intercept
summary(m3.2)

```



##  Finding covariates to explain variation
Why are some datasets affected while others aren't? Does it have something to do with sequence stats in the training data?
```{r extracting stat data}
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv"))
full_df <- merge(df_rel, stat_file, by = "Species")
```

```{r some stats}

df <- full_df %>% #*** Updated to remove "No"
  filter(smoothing_type != "No") %>%
  mutate(across(
    c(mean_transitions, total_transitions, Num_Activities),
    ~ scale(.x, center = TRUE, scale = TRUE)[,1]
  ))

null_m <- glmmTMB(relative_change ~ -1 + smoothing_type + (1|Species/rep_id), data = df)

m3 <- glmmTMB(relative_change ~ -1 + smoothing_type +
                    mean_transitions + multi_behaviour + Num_Activities +
                    (1|Species/rep_id), data = df)

anova(null_m, m3)
r.squaredGLMM(m3)

summary(m3)


# but maybe I should be looking at the interactions between those variables
m4 <- glmmTMB(relative_change ~ smoothing_type *
                (mean_transitions + multi_behaviour + Num_Activities) +
                (1|Species/rep_id), data = df)
anova(m3, m4)
summary(m4)

```
