---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5)
```

This is the report to compare the comparisons between the species datasets.

```{r bind, include = FALSE}
# Firstly, loading in the individual files and creating the collated performance file

# base_path <- "C:/Users/PC/OneDrive - University of the Sunshine Coast/PostProcessing"
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>%
  mutate(smoothing_type = gsub("Smoothing", "", smoothing_type))

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")) #, "LSTM"))
data$Species <- factor(data$Species, levels = c(unique(data$Species)))

# save this
#print(head(data %>% dplyr::filter(Metric == "F1")))
fwrite(data, file.path(base_path, "Output", "collated_performance.csv"))
```

## Base Performance
Looking at base performance.
```{r}
means <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1", smoothing_type == "No") %>%
  group_by(smoothing_type, Species) %>%
  summarise(Score = mean(Score))

print("best performance")
max(means$Score)
print("worst performance")
min(means$Score)

```

## Overall effect
Firstly I'll look at how much improvement there was overall (across all datasets)
```{r Stats on the overall effect}
df <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1")
# and also do it as relative from the control
# Get NoSmoothing scores by Species
baseline <- df %>%
  filter(smoothing_type == "No") %>%
  select(Species, rep_id, baseline = Score)

df_rel <- df %>%
  left_join(baseline, by = c("Species", "rep_id")) %>%
  mutate(relative_change = Score - baseline)

# do some stats here (note that No is set to the reference level)

# v1: what I thought was right
simple_model <- lmer(relative_change ~ smoothing_type + (1|Species/rep_id), data = df_rel)
summary(simple_model)
# yay significant results

# v2: what Dave suggested
dave_model <- lmer(relative_change ~ smoothing_type * Species + (1|rep_id), data = df_rel)
summary(dave_model)
# frighteningly... bayesian isnt significant anymore... oh no... this is bad news for me
```

```{r plots on the overall effect, fig.width = 8, fig.height = 8}
# Summarise mean and 95% CI
summary_df <- df_rel %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    n = n(),
    mean_rel = mean(relative_change, na.rm = TRUE),
    se = sd(relative_change, na.rm = TRUE) / sqrt(n),
    ci = 1.96 * se, # gives us the confidence intervals (i think)
    .groups = "drop"
  )
# this is 0 every time, so replace with a flat line for visually cleaner
ggplot(summary_df %>% dplyr::filter(!smoothing_type == "No"),
       aes(x = smoothing_type,
           y = mean_rel,
           colour = smoothing_type)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_rel - ci,
                    ymax = mean_rel + ci),
                width = 0.5) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  my_theme() +
  scale_colour_manual(values = my_colours) +
  labs(x = "Smoothing Type",
       y = "Mean change in F1 (relative to No smoothing)",
       fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  facet_wrap(~Species, ncol = 3)
```

## Within species
Breaking it down and looking at each species individually.
```{r stats on the individual species}
# per species
species <- unique(df_rel$Species)
all_stats <- data.frame()
for (x in species){
  dat <- df_rel %>% dplyr::filter(Species == x)
  mod <- lm(
    relative_change ~ smoothing_type + (1 | rep_id), # when I removed the 0+ the sigificant dissapeared 0_0
    data = dat
  )

  stats <- as.data.frame(summary(mod)$coefficients)
  stats$Species <- x
  stats$sig <- ifelse(stats$`Pr(>|t|)` < 0.05, TRUE, FALSE)
  stats$smoothing_type <- gsub("smoothing_type", "", rownames(stats))
  
  all_stats <- rbind(all_stats, stats)
}

fwrite(all_stats, file.path(base_path, "Output", "WithinSpeciesStats.csv"))

sig_stats <- all_stats %>% dplyr::filter(sig == TRUE)
sig_stats
```

##  Finding covariates to explain variation
Why are some datasets affected while others aren't? Does it have something to do with sequence stats in the training data?
```{r extracting stat data}
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv"))
full_df <- merge(df_rel, stat_file, by = "Species")
```

```{r some stats}
# scale everything so one variable doesnt have much bigger effect
full_df <- full_df %>%
  mutate(across(
    c(mean_transitions, total_transitions, Num_Activities),
    ~ scale(.x, center = TRUE, scale = TRUE)[,1]
  ))

full_model <- lmer(
  relative_change ~
    smoothing_type * Species *
      (mean_transitions +
       multi_behaviour +
       Num_Activities) +
    (1 | rep_id),
  data = full_df,
  REML = FALSE
)

model_null <- lmer(
  relative_change ~ smoothing_type +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)

anova(model_null, full_model)
summary(full_model)

library(MuMIn)
r.squaredGLMM(full_model)

```
