---
title: "ComparingComparisonsReport"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path: "C:/Users/oaw001/OneDrive - University of the Sunshine Coast/PostProcessing"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 8, fig.height = 5)
```

This is the report to compare the comparisons between the species datasets.

Step 1 is to pull together all the data from each species. 
```{r bind, include = FALSE}
files <- list.files(file.path(base_path, "Output"), pattern = "all_smoothed_metrics.csv", full.names = TRUE, recursive = TRUE)

data <- lapply(files, function(x){
  data <- fread(x)
  data$Species <- basename(dirname(x))
  return(data)
})

data <- bind_rows(data)
data <- data %>% dplyr::filter(Activity == "Macro-Average") %>%
  mutate(smoothing_type = gsub("Smoothing", "", smoothing_type))

# add levels so its easier to plot later
data$smoothing_type <- factor(data$smoothing_type, levels = c("No", "Mode", "Duration", "Transition", "HMM", "Bayesian")) #, "LSTM"))

data$Species <- factor(data$Species, levels = c(unique(data$Species)))

# save this
#print(head(data %>% dplyr::filter(Metric == "F1")))
fwrite(data, file.path(base_path, "Output", "collated_performance.csv"))
```

Step 2 is to compare these results. 
```{r compare}
# its hard to make any one single judgement, so I can make a plot for now

# Plot so that control is more obvious
df <- data %>% dplyr::filter(Activity == "Macro-Average", Metric == "F1")
ggplot(df, aes(x = smoothing_type, y = Score, fill = smoothing_type)) +
  geom_boxplot() +
  my_theme() +
  scale_fill_manual(values = my_colours) +
  labs(x = "Activity", y = "Score", fill = "Smoothing Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species)


# and also do it as relative from the control
setDT(df)

# Get NoSmoothing scores by Species
df[, rep_id := seq_len(.N), by = .(Species, smoothing_type)]
baseline <- df[
  smoothing_type == "No",
  .(Species, rep_id, baseline = Score)
]
df <- merge(
  df,
  baseline,
  by = c("Species", "rep_id"),
  all.x = TRUE
)
df[, relative_change := Score - baseline]

# now plot that
df_sum <- df %>%
  filter(smoothing_type != "No") %>%
  group_by(Species, smoothing_type) %>%
  summarise(
    mean_rc = mean(relative_change, na.rm = TRUE),
    se_rc   = sd(relative_change, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

ggplot(df_sum, aes(x = smoothing_type, y = mean_rc, colour = smoothing_type)) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "black") +
  geom_point(size = 2.5) +
  geom_errorbar(
    aes(ymin = mean_rc - se_rc, ymax = mean_rc + se_rc),
    width = 0.2
  ) +
  my_theme() +
  scale_colour_manual(values = my_colours) +
  labs(x = "Smoothing Type", y = "Relative change to F1 Score") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~Species, ncol = 5) +
  theme(legend.position = "none")

```

Look at the stats for that. Begin with very general summaries and descriptions.

```{r}
# add the 2 information sources together
print("best change")
max(df$relative_change)
print("worst change")
min(df$relative_change)

# mean changes
summaries <- df %>%
  group_by(Species, smoothing_type) %>%
  summarise(meanF1 = mean(Score),
            meanChange = mean(relative_change)) %>%
  arrange(desc(meanChange))
summaries

summaries %>% group_by(smoothing_type) %>%
  summarise(max = max(meanChange), mean = mean(meanChange))

```

Within species
```{r}
# per species
species <- unique(df$Species)
all_stats <- data.frame()
for (x in species){
  dat <- df %>% dplyr::filter(Species == x)
  dat$smoothing_type <- relevel(dat$smoothing_type, ref = "No")
  mod <- lmer(
    relative_change ~ 0 + smoothing_type + (1 | rep_id),
    data = dat
  )

  stats <- as.data.frame(summary(mod)$coefficients)
  stats$Species <- x
  stats$sig <- ifelse(stats$`Pr(>|t|)` < 0.05, TRUE, FALSE)
  stats$smoothing_type <- gsub("smoothing_type", "", rownames(stats))
  
  all_stats <- rbind(all_stats, stats)
}

fwrite(all_stats, file.path(base_path, "Output", "WithinSpeciesStats.csv"))

sig_stats <- all_stats %>% dplyr::filter(sig == TRUE)
sig_stats
```

Across species
```{r}
# all species
simple_model <- lmer(relative_change ~ smoothing_type + (1|Species/rep_id), data = df)
summary(simple_model)
```

Step 3 is to correlate this with the sequential stats I was able to draw out in the sequence report on the training data. Why are some datasets positively affected while others aren't?

```{r extracting stat data}
stat_file <- fread(file.path(base_path, "Output", "All_sequence_stats.csv"))

full_df <- merge(df, stat_file, by = "Species")
```

```{r some stats}

full_model <- lmer(
  relative_change ~ smoothing_type * Transition_Rate * Prop_Transitions * Num_Activities +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)
#summary(full_model)

simple_model <- lmer(
  relative_change ~ smoothing_type * Transition_Rate * Num_Activities +
    smoothing_type * Prop_Transitions * Num_Activities +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)
#summary(simple_model)

very_simple_model <- lmer(
  relative_change ~ smoothing_type +
    (1 | Species / rep_id),
  data = full_df, REML = FALSE
)

anova(full_model, simple_model, very_simple_model)

sig_terms <- summary(full_model)$coefficients %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(`Pr(>|t|)` < 0.05)
sig_terms

```
